A Type System for Safe and Scalable Parallelism using Logical Regions

### 2012-10-16 doc/types-pldi2013/langdef.tex

Our goal is to show $\privconsist{E}{\Phi}$ holds for any execution of an expression $e$ with static privileges $\Phi$.
As usual, even stating such a claim requires that corresponding parts of the initial static and dynamic environments be consistent.
For our results, three consistency properties are needed:
\begin{itemize}
\item $\mapconsist{\Omega}$, a region mapping satisfies the region constraints under which the expression was typed
\item $\localconsist{L}{\Gamma}$, local values have types consistent with the environment under which the expression was typed
\item $\storeconsist{S}{H}$, the initial state of the store contains values whose types are consistent with the heap typing
\end{itemize}

\noindent Two additional properties are proven for each subexpression:
\begin{itemize}
\item $\valueconsist{v}{T}$, the expression yields a value of the right type
\item $\traceconsist{E}$, values written to the store match the heap types of their locations
\end{itemize}
Figure~\ref{fig:constprop} gives the precise definition of these properties.

@@

We spare the reader the lengthy proof, which can be found in
\cite{LegionTypes12}, and merely outline the general strategy.  The
proof is a standard induction on the structure of the derivation in
the operational semantics.  The proof is organized into six pieces,
one for each of the three consistency properties that must be shown
for the execution environment of subexpressions and one for each of
the three properties of the execution's results we wish to show:

@@

\vspace{4pt}
\hangindent=2\parindent \hangafter=1 \noindent %
$\privconsist{E}{\Phi}$ -  The proof of the crucial property of containment of heap accesses within the available privileges
is similar in outline to the previous step.  The easy cases are covered by a lemma combining the corresponding results
for subexpressions: if $\privconsist{E_1}{\Phi}$ and $\privconsist{E_2}{\Phi}$ and $any\_interleave(E', E_1, E_2)$, then $\privconsist{E'}{\Phi}$.
Straightforward proofs cover {\tt read}, {\tt write}, {\tt reduce}, and one final special case for function calls.

%% \begin{table*}
%% \centering
%% {\small
%% \begin{tabular}{l|p{5cm}|p{5cm}|p{5cm}}
%% {\bf Rule} & 
%% \multicolumn{1}{|c|}{$\mathbf{v \sim_H M \llbracket T \rrbracket}$} &
%% \multicolumn{1}{|c|}{$\mathbf{E \sim H}$} &
%% \multicolumn{1}{|c|}{$\mathbf{E : \Phi}$} \\
%% \hline
%% BoolConst & & & \\
%% IntConst & & & \\
%% TupleBuild & & & \\
%% TupleField & & & \\
%% LocalVar & & & \\
%% NullPtr & & & \\
%% NewPtr & & & \\
%% IsNull & & & \\
%% UpRegion & & & \\
%% DownRegion & & & \\
%% Read & & & \\
%% Write & & & \\
%% Reduce & & & \\
%% Newcolor & & & \\
%% Color & & & \\
%% IntMath & & & \\
%% IntCmp & & & \\
%% Let & & & \\
%% IfElse & & & \\
%% TaskCall & & & \\
%% Partition & & & \\
%% Pack & & & \\
%% Unpack & & &
%% \end{tabular}}
%% \caption{Proof pieces}
%% \end{table*}

>>>>

Our goal is to show $\privconsist{E}{\Phi}$ holds for any execution of an expression $e$ with static privileges $\Phi$.
As usual, even stating such a claim requires that corresponding parts of the initial static and dynamic environments be consistent.
For our results, three consistency properties are needed:
\begin{itemize}
\item mapping consistency, written $\mapconsist{\Omega}$, guarantees a region mapping $M$ satisfies the region constraints $\Omega$ under
which an expression was typed
\item local value consistency, written $\localconsist{L}{\Gamma}$, guarantees local values in $L$ have types consistent with the environment under which the expression was typed
\item store consistency, written $\storeconsist{S}{H}$, guarantees every location in the store $S$ contains a value consistent with the location's type in the heap typing $H$
\end{itemize}

\noindent Two additional properties are proven for each subexpression:
\begin{itemize}
\item result value consistency, written $\valueconsist{v}{T}$, guarantees any evaluation of an expression yields a value of the right type
\item memory trace consistency, written $\traceconsist{E}$, guarantees that all writes and reductions use values of the right types
\end{itemize}
Figure~\ref{fig:constprop} gives the precise definition of these properties.

@@

The proof, which can be found in \cite{LegionTypes12}, is lengthy but
straightforward due to the way in which the operational semantics were
constructed.  It proceeds by standard induction on the structure of the
derivation in the operational semantics, at each step showing that
subexpressions have consistent starting environments and that the
value and trace consistency of the subexpressions is sufficient to prove
the consistency of the parent expression.

\texcomment{
We spare the reader the lengthy proof, which can be found in
\cite{LegionTypes12}, and merely outline the general strategy.  The
proof is a standard induction on the structure of the derivation in
the operational semantics.  The proof is organized into six pieces,
one for each of the three consistency properties that must be shown
for the execution environment of subexpressions and one for each of
the three properties of the execution's results we wish to show:

@@

\vspace{4pt}
\hangindent=2\parindent \hangafter=1 \noindent %
$\privconsist{E}{\Phi}$ -  The proof of the crucial property of containment of heap accesses within the available privileges
is similar in outline to the previous step.  The easy cases are covered by a lemma combining the corresponding results
for subexpressions: if $\privconsist{E_1}{\Phi}$ and $\privconsist{E_2}{\Phi}$ and $any\_interleave(E', E_1, E_2)$, then $\privconsist{E'}{\Phi}$.
Straightforward proofs cover {\tt read}, {\tt write}, {\tt reduce}, and one final special case for function calls. 
}
